How to compute the fusion product of MV cycles in type A

### 2021-04-13 draft.tex
Notation for examples.

We view MV cycles as semi-standard young tableaux. Suppose we have two tableaux $\tau_1$ and $\tau_2$ with respective weights $\lambda_i$ and $\mu_i$ and we wished to take their fusion product. We consider a generic matrix $X \in \UU^{\mu_1,\mu_2}_{0,s}$ and see what relations on the variables in $X$ occur when also asking for it to be in $\OO^{\lambda_1,\lambda_2}_{0,s}$.

Consider the subtableaux $\tau_{1,i}$ of $\tau_1$ and $\tau_{2,i}$ of $\tau_2$ where we take the boxes of $\tau_1$ and $\tau_2$ containing the numbers $1,\dots,i$. Let $\lambda_{1,j}$ and $\lambda_{2,j}$ be the corresponding sizes and $\mu_{1,j}$ and $\mu_{2,j}$ be the corresponding weights of the subtableaux. For each $i$, we obtain a corresponding submatrix $X_i \in \UU^{\mu_{1,i}, \mu_{2,i}}_{0,s}$ of $X$. These subtableaux impose rank conditions, equivalently vanishing minors, on each $X_i$ when asking them to be in $\OO^{\lambda_{1,i}, \lambda_{2,i}}_{0,s}$, which allows us to deduce relations on the variables in $X$. The relations must also include the new variables in the latest block and cannot have $s$ as a factor.

@@

\begin{example}
Let $G = \GL_3$ and consider the tableaux
$$\tau_1 = \young(12,3) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2=\young(13,2).$$

Lusztig data: $(010)$ for $\tau_1$, $(101)$ for $\tau_2$

@@

The MV cycles corresponding to $\tau_1$ and $\tau_2$ are both isomorphic to $\PP^2$.
The weights are $\lambda_1 = \lambda_2 = (2,1,0)$ and $\mu_1 = \mu_2 = (1,1,1)$. Hence the matrix in consideration is 
\[
X = \left[\begin{BMAT}(e){cc;cc;cc}{cc;cc;cc}
    0 & 1 & & & & \\
     & s & a_1 & a_2 & b_1 & b_2 \\
     & & 0 & 1 & & \\
     & & & s & b_3 & b_4 \\
     & & & & 0 & 1 \\
     & & & & & s
\end{BMAT}
\right].
\]

For the first entry of each $\mu_i$, we have the following subtableaux to consider:
$$\young(1) \hspace{.5cm} \text{ and } \hspace{.5cm} \young(1).$$
Let $X_1$ be the top-left $2 \times 2$ submatrix of $X$. The subtableaux imposes the conditions $\dim \ker X_1 = 1$ and $\dim \ker (X_1-s) = 1$. We see that we do not obtain any relations from this.

For the first two entries of each $\mu_i$, we have the following subtableaux:
$$\young(12) \hspace{.5cm} \text{ and } \hspace{.5cm} \young(1,2).$$
Let $X_2$ be the top-left $4 \times 4$ submatrix of $X$. The left subtableau imposes the conditions $\dim \ker X_2 = 1$ and $\dim \ker X_2^2 = 2$ while the right subtableau imposes the condition that $\dim \ker (X_2 -s) = 2$. These conditions force the relation $$a_1 + sa_2 = 0.$$

The original tableaux for $X$ imposes the conditions that $\dim \ker X = 2$, $\dim \ker X^2 = 3$, $\dim \ker (X-s) = 2$, and $\dim \ker (X-s)^2 = 3$. These force the relation $$b_3 = 0.$$

Hence we consider the quotient
$$\frac{\CC[a_1,a_2,b_1,b_2,b_3,b_4,s]}{\langle a_1+sa_2,b_3,s \rangle} \cong
\frac{\CC[a_1,a_2,b_1,b_2,b_3,b_4]}{\langle a_1,b_3 \rangle}.$$
We see already that the $0$-fibre is irreducible. The tableau corresponding to the ideal $\langle a_1, b_3 \rangle$ is 
$$\young(1123,23)$$
which corresponds to the MV cycle that is the product of the MV cycles corresponding to $\tau_1$ and $\tau_2$.

@@

\begin{example}
Let $G = SL_3$ and consider the tableaux
$$\tau_1 = \young(12) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2 = \young(11,23).$$

Lustig data: $(100)$ for $\tau_1$, $(001)$ for $\tau_2$.

Dimensions: $(1,-1,0)$ for $\tau_1$, $(0,1,-1)$ for $\tau_2$.

The MV cycles corresponding to the above tableaux are each isomorphic to $\PP^1$.
We have the weights to be $\lambda_1 = (2,0,0)$, $\lambda_2 = (2,2,0)$, $\mu_1 = (1,1,0)$, and $\mu_2 = (2,1,1)$. The matrix under consideration is therefore
\[
X = \left[\begin{BMAT}(e){ccc;cc;c}{ccc;cc;c}
    0 & 1 & & & & \\
     & 0 & 1 & & & \\
     & -s^2 & 2s & a_1 & a_2 & b_1 \\
     & & & 0 & 1 & \\
     & & & & s & b_2 \\
     & & & & & s
\end{BMAT}
\right].
\]
There are no relations from looking at the submatrix $X_1$. From the submatrix $X_2$, we require $$a_1 + sa_2 = 0.$$
From $X$, we further obtain the relations $$a_2b_2 + sb_1 = 0, a_1b_2 + s^2b_1 = 0.$$
Then the quotient ring we find is
$$\frac{\CC[a_1,a_2,b_1,b_2,s]}{\langle a_1+sa_2, a_2b_2+sb_1, a_1b_2+s^2b_1, s\rangle} \cong \frac{\CC[a_1,a_2,b_1,b_3,s]}{\langle a_1, a_2\rangle \cap \langle a_1, b_2 \rangle}.$$
We see that the 0-fibre is reducible with two components. The tableau corresponding to each ideal is
$$\begin{array}{cccc} \vspace{1mm}
    \young(1113,22) & \text{for } \langle a_1, a_2 \rangle & \text{Lusztig: }(101) & \dim = (1,0,-1)\\ 
    \young(1112,23) & \text{for } \langle a_1, b_2 \rangle & \text{Lusztig: }(010) & \dim = (1,0,-1).
\end{array}$$
Each of these tableaux corresponds to a MV cycle isomorphic to $\PP^2$ so the generic fibre $\PP^1 \times \PP^1$ degenerates to two copies of $\PP^2$.
\end{example}

\begin{example}
Let $G = SL_3$ and consider the following tableaux:
$$\tau_1 = \young(1122) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2 = \young(1111,2233).$$

Lusztig data: $(200)$ for $\tau_1$, $(002)$ for $\tau_2$.

Dimensions: $(2,-2,0)$ for $\tau_1$, $(0,2,-2)$ for $\tau_2$.

The MV cycles corresponding to these tableaux are each isomorphic to $\PP^1 \times \PP^1$. 
The weights are then $\lambda_1 = (4,0,0)$, $\lambda_2 = (4,4,0)$, $\mu_1 = (2,2,0)$, and $\mu_2 = (4,2,2)$.
Then we are considering the following matrix:
\[
X = \left[\begin{BMAT}(e){cccccc;cccc;cc}{cccccc;cccc;cc}
    0 & 1 & & & & & & & & & & \\
     & 0 & 1 & & & & & & & & & \\
     & & 0 & 1 & & & & & & & & \\
     & & & 0 & 1 & & & & & & & \\
     & & & & 0 & 1 & & & & & & \\
     & & -s^4 & 3s^3 & -6s^2 & 4s & a_1 & a_2 & a_3 & a_4 & b_1 & b_2 \\
     & & & & & & 0 & 1 & & & & \\
     & & & & & & & 0 & 1 & & & \\
     & & & & & & & & 0 & 1 & & \\
     & & & & & & & & -s^2 & 2s & b_3 & b_4 \\
     & & & & & & & & & & 0 & 1 \\
     & & & & & & & & & & -s^2 & 2s
\end{BMAT}
\right].
\]
By computer, it can be checked that we obtain the following 3 ideals:
$$
\begin{array}{cl}
    I_1 = & \langle a_1, a_2, a_3, a_4, s \rangle \\
    I_2 = & \langle a_1, a_2, b_3, b_4, s \rangle \\
    I_3 = & \langle a_1, a_2, b_3^2, a_4b_3 + a_3b_4, a_3b_3, a_3^2, s \rangle.
\end{array}
$$
It can be checked that $I_3$ has multiplicity 2 \rcom{Not sure how to make this precise and how to check it.} so the 0-fibre has 4 components, two of which are the same. The tableau corresponding to each ideal is
$$
\begin{array}{cccc}\vspace{1mm}
    \young(11111133,2222) & \text{for } I_1 & \text{Lusztig: }(202) & \dim = (2,0,-2) \\ \vspace{1mm}
    \young(11111122,2233) & \text{for } I_2 & \text{Lusztig: }(020) & \dim = (2,0,-2) \\ \vspace{1mm}
    \young(11111123,2223) & \text{for } I_3 & \text{Lusztig: }(111) & \dim = (2,0,-2).
\end{array}
$$
The MV cycle for each ideal is isomorphic to $\PP^2 \times \PP^2$.
\end{example}

\begin{example}
Let us redo example 3 but with non-dominant weights $\mu_i$ such that $\mu_1 + \mu_2$ is still dominant. Then consider the tableaux
$$\tau_1 = \young(22) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2 = \young(11,33).$$
Then the weights are $\lambda_1 = (2,0,0)$, $\lambda_2 = (2,2,0)$, $\mu_1 = (0,2,0)$, and $\mu_2 = (2,0,2)$. Our matrix $X$ is now
\[
X = \left[\begin{BMAT}(e){cc;cc;cc}{cc;cc;cc}
    0 & 1 & & & & \\
    -s^2 & 2s & a_1 & a_2 & b_1 & b_2 \\
     & & 0 & 1 & & \\
     & & & 0 & b_3 & b_4 \\
     & & & & 0 & 1 \\
     & & & & -s^2 & 2s
\end{BMAT}
\right].
\]
Using a computer, the ideals we now get are
$$
\begin{array}{cl}
    I_1 = & \langle a_1, a_2 \rangle \\
    I_2 = & \langle b_3, b_4 \rangle \\
    I_3 = & \langle b_3^2, a_2b_3 + a_1b_4, a_1b_3, a_1^2 \rangle. 
\end{array}
$$
The tableaux corresponding to each ideal is
$$
\begin{array}{cc}\vspace{1mm}
    \young(1133,22) & \text{for } I_1 \\ \vspace{1mm}
    \young(1122,33) & \text{for } I_2 \\ \vspace{1mm}
    \young(1123,23) & \text{for } I_3
\end{array}
$$
and the MV cycles corresponding to each are the same as before.
\end{example}

>>>>

We view MV cycles as semi-standard young tableaux. Suppose we have two tableaux $\tau'$ and $\tau''$, with respective weights $\lambda'$, $\mu'$ and $\lambda''$, $\mu''$, and we wished to take their fusion product. We consider a generic matrix $X \in \UU^{\mu',\mu''}_{0,s}$ and see what relations on the variables in $X$ occur when also asking for it to be in $\OO^{\lambda',\lambda''}_{0,s}$.

Consider the subtableaux $\tau'_{i}$ of $\tau'$ and $\tau''_{i}$ of $\tau''$ where we take the boxes of $\tau'$ and $\tau''$ containing the numbers $1,\dots,i$. Let $\lambda'_{j}$ and $\lambda''_{j}$ be the corresponding sizes and $\mu'_{j}$ and $\mu'_{j}$ be the corresponding weights of the subtableaux. For each $i$, we obtain a corresponding submatrix $X_i \in \UU^{\mu'_{i}, \mu''_{i}}_{0,s}$ of $X$. These subtableaux impose rank conditions, equivalently vanishing minors, on each $X_i$ when asking them to be in $\OO^{\lambda'_{i}, \lambda''_{i}}_{0,s}$, which allows us to deduce relations on the variables in $X$. The relations must also include the new variables in the latest block and cannot have $s$ as a factor.

@@

\begin{example}
Let $G = \GL_3$ and consider the tableaux
$$\tau' = \young(12,3) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau''=\young(13,2).$$

Lusztig data: $(010)$ for $\tau'$, $(101)$ for $\tau''$

@@

The MV cycles corresponding to $\tau'$ and $\tau''$ are both isomorphic to $\PP^2$.
The weights are $\lambda' = \lambda'' = (2,1,0)$ and $\mu' = \mu'' = (1,1,1)$. Hence the matrix in consideration is 
\[
X = \left[\begin{BMAT}(e){cc;cc;cc}{cc;cc;cc}
    0 & 1 & & & & \\
     & s & A_{12}^1 & A_{12}^2 & A_{13}^1 & A_{13}^2 \\
     & & 0 & 1 & & \\
     & & & s & A_{23}^1 & A_{23}^2 \\
     & & & & 0 & 1 \\
     & & & & & s
\end{BMAT}
\right].
\]

For the first entry of $\mu'$ and $\mu''$, we have the following subtableaux to consider:
$$\young(1) \hspace{.5cm} \text{ and } \hspace{.5cm} \young(1).$$
Let $X_1$ be the top-left $2 \times 2$ submatrix of $X$. The subtableaux imposes the conditions $\dim \ker X_1 = 1$ and $\dim \ker (X_1-s) = 1$. We see that we do not obtain any relations from this.

For the first two entries of $\mu'$ and $\mu''$, we have the following subtableaux:
$$\young(12) \hspace{.5cm} \text{ and } \hspace{.5cm} \young(1,2).$$
Let $X_2$ be the top-left $4 \times 4$ submatrix of $X$. The left subtableau imposes the conditions $\dim \ker X_2 = 1$ and $\dim \ker X_2^2 = 2$ while the right subtableau imposes the condition that $\dim \ker (X_2 -s) = 2$. These conditions force the relation $$A_{12}^1 + sA_{12}^2 = 0.$$

The original tableaux for $X$ imposes the conditions that $\dim \ker X = 2$, $\dim \ker X^2 = 3$, $\dim \ker (X-s) = 2$, and $\dim \ker (X-s)^2 = 3$. These force the relation $$A_{23}^1 = 0.$$

Hence we consider the quotient
$$\frac{\CC[A_{12}^1,A_{12}^2,A_{13}^1,A_{13}^2,A_{23}^1,A_{23}^2,s]}{\langle A_{12}^1+sA_{12}^2,A_{23}^1,s \rangle} \cong
\frac{\CC[A_{12}^1,A_{12}^2,A_{13}^1,A_{13}^2,A_{23}^1,A_{23}^2]}{\langle A_{12}^1,A_{23}^1 \rangle}.$$
We see already that the $0$-fibre is irreducible. The tableau corresponding to the ideal $\langle A_{12}^1,A_{23}^1 \rangle$ is 
$$\young(1123,23)$$
which corresponds to the MV cycle that is the product of the MV cycles corresponding to $\tau'$ and $\tau''$.

@@

\begin{example}
Let $G = SL_3$ and consider the tableaux
$$\tau' = \young(12) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau'' = \young(11,23).$$

Lustig data: $(100)$ for $\tau'$, $(001)$ for $\tau''$.

Dimensions: $(1,-1,0)$ for $\tau'$, $(0,1,-1)$ for $\tau''$.

The MV cycles corresponding to the above tableaux are each isomorphic to $\PP^1$.
We have the weights to be $\lambda' = (2,0,0)$, $\lambda'' = (2,2,0)$, $\mu' = (1,1,0)$, and $\mu'' = (2,1,1)$. The matrix under consideration is therefore
\[
X = \left[\begin{BMAT}(e){ccc;cc;c}{ccc;cc;c}
    0 & 1 & & & & \\
     & 0 & 1 & & & \\
     & -s^2 & 2s & A_{12}^1 & A_{12}^2 & A_{13}^1 \\
     & & & 0 & 1 & \\
     & & & & s & A_{23}^1 \\
     & & & & & s
\end{BMAT}
\right].
\]
There are no relations from looking at the submatrix $X_1$. From the submatrix $X_2$, we require $$A_{12}^1 + sA_{12}^2 = 0.$$
From $X$, we further obtain the relations $$A_{12}^2A_{23}^1 + sA_{13}^1 = 0, A_{12}^1A_{23}^1 + s^2A_{13}^1 = 0.$$
Then the quotient ring we find is
$$\frac{\CC[A_{12}^1,A_{12}^2,A_{13}^1,A_{23}^1,s]}{\langle A_{12}^1+sA_{12}^2, A_{12}^2A_{23}^1+sA_{13}^1, A_{12}^1A_{23}^1+s^2A_{13}^1, s\rangle} \cong \frac{\CC[A_{12}^1,A_{12}^2,A_{13}^1,b_3,s]}{\langle A_{12}^1, A_{12}^2\rangle \cap \langle A_{12}^1, A_{23}^1 \rangle}.$$
We see that the 0-fibre is reducible with two components. The tableau corresponding to each ideal is
$$\begin{array}{cccc} \vspace{1mm}
    \young(1113,22) & \text{for } \langle A_{12}^1, A_{12}^2 \rangle & \text{Lusztig: }(101) & \dim = (1,0,-1)\\ 
    \young(1112,23) & \text{for } \langle A_{12}^1, A_{23}^1 \rangle & \text{Lusztig: }(010) & \dim = (1,0,-1).
\end{array}$$
Each of these tableaux corresponds to a MV cycle isomorphic to $\PP^2$ so the generic fibre $\PP^1 \times \PP^1$ degenerates to two copies of $\PP^2$.
\end{example}

\begin{example}
Let $G = SL_3$ and consider the following tableaux:
$$\tau' = \young(1122) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau'' = \young(1111,2233).$$

Lusztig data: $(200)$ for $\tau'$, $(002)$ for $\tau''$.

Dimensions: $(2,-2,0)$ for $\tau'$, $(0,2,-2)$ for $\tau''$.

The MV cycles corresponding to these tableaux are each isomorphic to $\PP^1 \times \PP^1$. 
The weights are then $\lambda' = (4,0,0)$, $\lambda'' = (4,4,0)$, $\mu' = (2,2,0)$, and $\mu'' = (4,2,2)$.
Then we are considering the following matrix:
\[
X = \left[\begin{BMAT}(e){cccccc;cccc;cc}{cccccc;cccc;cc}
    0 & 1 & & & & & & & & & & \\
     & 0 & 1 & & & & & & & & & \\
     & & 0 & 1 & & & & & & & & \\
     & & & 0 & 1 & & & & & & & \\
     & & & & 0 & 1 & & & & & & \\
     & & -s^4 & 3s^3 & -6s^2 & 4s & A_{12}^1 & A_{12}^2 & A_{12}^3 & A_{12}^4 & A_{13}^1 & A_{13}^2 \\
     & & & & & & 0 & 1 & & & & \\
     & & & & & & & 0 & 1 & & & \\
     & & & & & & & & 0 & 1 & & \\
     & & & & & & & & -s^2 & 2s & A_{23}^1 & A_{23}^2 \\
     & & & & & & & & & & 0 & 1 \\
     & & & & & & & & & & -s^2 & 2s
\end{BMAT}
\right].
\]
By computer, it can be checked that we obtain the following 3 ideals:
$$
\begin{array}{cl}
    I_1 = & \langle A_{12}^1, A_{12}^2, A_{12}^3, A_{12}^4, s \rangle \\
    I_2 = & \langle A_{12}^1, A_{12}^2, A_{23}^1, A_{23}^2, s \rangle \\
    I_3 = & \langle A_{12}^1, A_{12}^2, (A_{23}^1)^2, A_{12}^4A_{23}^1 + A_{12}^3A_{23}^2, A_{12}^3A_{23}^1, (A_{12}^3)^2, s \rangle.
\end{array}
$$
It can be checked that $I_3$ has multiplicity 2 \rcom{Not sure how to make this precise and how to check it.} so the 0-fibre has 4 components, two of which are the same. The tableau corresponding to each ideal is
$$
\begin{array}{cccc}\vspace{1mm}
    \young(11111133,2222) & \text{for } I_1 & \text{Lusztig: }(202) & \dim = (2,0,-2) \\ \vspace{1mm}
    \young(11111122,2233) & \text{for } I_2 & \text{Lusztig: }(020) & \dim = (2,0,-2) \\ \vspace{1mm}
    \young(11111123,2223) & \text{for } I_3 & \text{Lusztig: }(111) & \dim = (2,0,-2).
\end{array}
$$
The MV cycle for each ideal is isomorphic to $\PP^2 \times \PP^2$.
\end{example}

\begin{example}
Let us redo example 3 but with non-dominant weights $\mu'$ and $\mu''$ such that $\mu' + \mu''$ is effective dominant. Consider the tableaux
$$\tau' = \young(22) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau'' = \young(11,33).$$
Then the weights are $\lambda' = (2,0,0)$, $\lambda'' = (2,2,0)$, $\mu' = (0,2,0)$, and $\mu'' = (2,0,2)$. Our matrix $X$ is now
\[
X = \left[\begin{BMAT}(e){cc;cc;cc}{cc;cc;cc}
    0 & 1 & & & & \\
    -s^2 & 2s & A_{12}^1 & A_{12}^2 & A_{13}^1 & A_{13}^2 \\
     & & 0 & 1 & & \\
     & & & 0 & A_{23}^1 & A_{23}^2 \\
     & & & & 0 & 1 \\
     & & & & -s^2 & 2s
\end{BMAT}
\right].
\]
Using a computer, the ideals we now get are
$$
\begin{array}{cl}
    I_1 = & \langle A_{12}^1, A_{12}^2 \rangle \\
    I_2 = & \langle A_{23}^1, A_{23}^2 \rangle \\
    I_3 = & \langle (A_{23}^1)^2, A_{12}^2A_{23}^1 + A_{12}^1A_{23}^2, A_{12}^1A_{23}^1, (A_{12}^1)^2 \rangle. 
\end{array}
$$
The tableaux corresponding to each ideal is
$$
\begin{array}{cc}\vspace{1mm}
    \young(1133,22) & \text{for } I_1 \\ \vspace{1mm}
    \young(1122,33) & \text{for } I_2 \\ \vspace{1mm}
    \young(1123,23) & \text{for } I_3
\end{array}
$$
and the MV cycles corresponding to each are the same as before.
\end{example}
